---
title: "GRaD Presentation"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

# Bookkeeping

```{r Packages, include = F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
library("emmeans")
library("formattable")
library("dplyr")
library("descr")
library("MASS")
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

```{r Model Labels}
labs.prop<-c(
  '(Intercept)' = "Intercept"
  , BEHAVIORHU = "Behavior"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  )

labs.bout.1<-c(
  '(Intercept)' = "Intercept"
  , BEHAVIORHU = "Behavior"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  , 'BEHAVIORHU:GENERALIZED_ENVIRONMENTGreen Area' = "Behavior: Generalized Environment"
)

labs.bout.2<-c(
  '(Intercept)' = "Intercept"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  )

labs.peck<-c(
  '(Intercept)' = "Intercept"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  , 'GENERALIZED_ENVIRONMENTGreen Area:DISTURBANCE_FREQUENCY' = "Generalized Environment: Disturbance Frequency"
  )
```

```{r DATA.SR & BOUT}
DATA.SR<-read.csv("DATA.SR.csv", stringsAsFactors = T) %>%
  rename("VIDEO_ID" = "VIDEO_ID."
         , "ID" = "ID.")
BOUT.raw<-read.csv("BOUT.csv", stringsAsFactors = T)
```

# Proportion Data

```{r Open PROP}
PROP<-DATA.SR[,c(1,2,15,17,19,22,30,36,41,46)]%>% 
  subset(.
         , HU_BEHAVIOR_PROPORTION_... != 0 
         & HD_BEHAVIOR_PROPORTION_... != 0 
         & M_BEHAVIOR_PROPORTION_... != 0
         ) %>% #Remove cases where proportion = 1 or 0
  rename("HU" = "HU_BEHAVIOR_PROPORTION_..."
         , "HD" = "HD_BEHAVIOR_PROPORTION_..."
         , "M" = "M_BEHAVIOR_PROPORTION_..."
         , "DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES"
         ) %>%
  pivot_longer(., cols = c("HU"
                              , "HD"
                              , "M")
               , names_to = "BEHAVIOR"
               , values_to = "PROPORTION")
PROP$ASIN_PROPORTION<-asin(sqrt(PROP$PROPORTION)) #Transform for normality

str(PROP)
```

```{r PROP Summary}
PROP.SUMMARY <- summarySE(data = PROP
                      , measurevar = "PROPORTION"
                      , groupvars = c("GENERALIZED_ENVIRONMENT"
                                      , "BEHAVIOR"
                                      )
                      )
PROP.SUMMARY
```

## Stacked barplot

```{r PROP Barplot}
PROP.BARPLOT<-ggplot(PROP.SUMMARY
                 , aes(x = GENERALIZED_ENVIRONMENT
                       , y = PROPORTION
                       , fill = BEHAVIOR))+
  geom_bar(stat = 'identity'
           , position = 'stack')+
  geom_text(aes(label = paste0(formattable::digits(PROPORTION*100, dig=2)
                               , "%"))
            , position = position_stack(vjust = 0.5)
            , size = 4) +
  scale_y_continuous(labels = scales::percent) +
  theme_classic()+
  ylab("Proportion of time")+
  scale_x_discrete(labels=c("Commercial Area"
                            , "Green Area"))+
  scale_fill_manual(values = cbPalette, labels = c("Foraging"
                                                   , "Alert"
                                                   , "Moving")
                    , name="")+
  theme(legend.position = "bottom"
        , text = element_text(size = 18)
        , axis.title.x = element_blank())
PROP.BARPLOT
```

## PROP Model

```{r PROP Model}
PROP.MOD<-lm(ASIN_PROPORTION~BEHAVIOR+GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+BAIT_PRESENCE+GROUP_SIZE, data = PROP)

sjPlot::tab_model(PROP.MOD
                  , pred.labels = labs.prop
                  , rm.terms = "BEHAVIORM"
                  , show.re.var = T
                  , title = "Proportion Model 1 Output"
                  , dv.labels = " Effects of on the proportion of behaviors")
```

Cool, nothing is significant (apart from "Moving" being different, but thats to be expected). 

Crows allocate similar time to foraging and vigilance.

I tinkered with the models, and the model with the lowest AIC value had no interactions and no random effects.

# Bout Data

```{r Open BOUT}
BOUT<-BOUT.raw %>%
  filter(.
         , DURATION > 0.01) %>% #Remove impossibly small values
  filter(.
         , BEHAVIOR != "M") #Remove "M"
BOUT$LDURATION<-log(BOUT$DURATION) #Transform data for normality
str(BOUT)
str(BOUT.raw)
```

```{r BOUT Summary}
BOUT.SUMMARY <- summarySE(data = BOUT
                      , measurevar = "DURATION"
                      , groupvars = c("GENERALIZED_ENVIRONMENT"
                                      , "BEHAVIOR"
                                      )
                      ) %>%
  mutate(BEHAVIOR = recode_factor(BEHAVIOR
                                  ,'HD' = 'Foraging'#Rename HD as 'foraging'
                                  , 'HU' = 'Alert') #Rename HU as 'Alert'
                           )

BOUT.SUMMARY
```

## BOUT Dot Plot

```{r BOUT Dot Plot}
BOUT.DOTPLOT<-BOUT.SUMMARY %>%
           ggplot(.
               , aes(x = GENERALIZED_ENVIRONMENT
                     , y = DURATION
                     , colour = GENERALIZED_ENVIRONMENT))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  geom_text(aes(label = round(DURATION, 2)), hjust = -0.2, size = 6) +
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , guide="none") +
  scale_x_discrete(labels = c("Commercial Area", "Green Area"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 18)) +
  facet_grid(~BEHAVIOR)

BOUT.DOTPLOT
```

> The error bars represent the standard error, while the symbols are the mean duration of bouts.

## BOUT Model - All behaviors

```{r BOUT Model - All behaviors}
BOUT.MOD<-rlmer(LDURATION~BEHAVIOR*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)

sjPlot::tab_model(BOUT.MOD
                  , pred.labels = labs.bout.1
                  , show.re.var = T
                  , title = "Bout Robust Model  Output"
                  , dv.labels = " Effects on duration of bouts of all behaviors")

sjPlot::plot_model(BOUT.MOD
                  , axis.labels = labs.bout.1
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on duration of bouts of all behaviors"
                  )
```

##BOUT model - Head Down

```{r BOUT Model - Head Down}
BOUT.MOD.HD<-rlmer(LDURATION~GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = subset(BOUT, BEHAVIOR == "HD"))

sjPlot::tab_model(BOUT.MOD.HD
                  , pred.labels = labs.bout.2
                  , show.re.var = T
                  , title = "Head Down"
                  , dv.labels = " Effects on foraging bout duration")

sjPlot::plot_model(BOUT.MOD.HD
                  , axis.labels = labs.bout.2
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on foraging bout duration")

```

```{r BOUT Model - Head Up}
BOUT.MOD.HU<-rlmer(LDURATION~GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = subset(BOUT, BEHAVIOR == "HU"))


sjPlot::tab_model(BOUT.MOD.HU
                  , pred.labels = labs.bout.2
                  , show.re.var = T
                  , title = "Head Up"
                  , dv.labels = " Effects on alert bout duration")

sjPlot::plot_model(BOUT.MOD.HU
                  , axis.labels = labs.bout.2
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on alert bout duration")

```

##BOUT Post Hoc

```{r BOUT Post Hoc}
BOUT.DIFF<-emmeans(BOUT.MOD, ~BEHAVIOR+GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
test(pairs(BOUT.DIFF, by= "BEHAVIOR"), adjust="fdr")
```

#Peck Rate

```{r Open PECK}
PECK<-DATA.SR[,c(1,2,15,17,19,22,30,38,48)] %>%
  subset(., HD_BEHAVIOR_DURATION > 0) %>%
  rename("DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES")
  
str(PECK)
```
Next, we'll compile the means of the peck rate by sentinel presence and generalized environment.

```{r PECK Summary, echo = F}
PECK.SUMMARY<-summarySE(PECK, "PECK_RATE"
                     , "GENERALIZED_ENVIRONMENT"
                     , na.rm=T)
PECK.SUMMARY
```

## PECK Dot Plot

```{r PECK Dot Plot}
PECK.DOTPLOT<-ggplot(data = PECK.SUMMARY
               , aes(x = GENERALIZED_ENVIRONMENT
                     , y = PECK_RATE
                     , color = GENERALIZED_ENVIRONMENT))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(PECK_RATE-se)
                    , ymax=(PECK_RATE+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  geom_text(aes(label = round(PECK_RATE, 2)), hjust = -0.2, size = 6) +
  theme_classic() +
  ylab("Mean Peck Rate (per min)") +
  scale_colour_manual(values = cbPalette
                      , guide="none") +
  scale_x_discrete(labels = c("Commercial Area"
                                  , "Green Area")) +
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 18))

PECK.DOTPLOT
```

## PECK Model

```{r PECK Model}
PECK.MOD<-rlmer(PECK_RATE~GENERALIZED_ENVIRONMENT*DISTURBANCE_FREQUENCY+GROUP_SIZE+BAIT_PRESENCE+(1|VIDEO_ID/ID), data = PECK)

sjPlot::tab_model(PECK.MOD
                  , pred.labels = labs.peck
                  , show.re.var = T
                  , title = "Robust Peck Rate Model Output"
                  , dv.labels = " Effects on peck rate")
sjPlot::plot_model(PECK.MOD
                  , axis.labels = labs.peck
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on peck rate" )
```

## Disturbance Frequency x Generalized Environment

```{r PECK Disturbance plot}
PECK.DISTURBANCE.DOTPLOT<-PECK %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
               , y = PECK_RATE
               , color = GENERALIZED_ENVIRONMENT))+
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = F) +
  theme_classic() +
  labs(y="Peck Rate (per min)", x="Disturbance Frequency (per min)")+
  scale_x_continuous(n.breaks=14)+
  theme(legend.position = "bottom"
        , text = element_text(size = 18))+
  scale_color_manual(values = cbPalette
                    , labels = c("Commercial Area"
                                 , "Green Area")
                    , name="")
  

PECK.DISTURBANCE.DOTPLOT
```

# Sentinel Likelihood

```{r open SENT}
SENT<-DATA.SR[1:81,c(1,7,10,15,17,19,20,22,30,31)] %>%
  distinct()
str(SENT)
```

```{r SENT Summary}
SENT.SUMMARY <- summarySE(data = SENT
                      , measurevar = "GROUP_SIZE" #Random
                      , groupvars = c("GENERALIZED_ENVIRONMENT"
                                      , "SENTINEL_PRESENCE"
                                      )
                      )

SENT.SUMMARY
```

##SENT Barplot

```{r SENT Barplot}
SENT.BARPLOT <- ggplot(SENT.SUMMARY, aes(x=GENERALIZED_ENVIRONMENT
                 , fill = SENTINEL_PRESENCE
                 , y = N
                 , group = SENTINEL_PRESENCE)) + 
  geom_bar(stat='identity', width=.5, position = "dodge") +
  geom_text(aes(label = N, group = SENTINEL_PRESENCE), position = position_dodge(width = 0.5), vjust = 1.5, size = 8) +
  ylab("Number of observations") +
  theme_classic() +
  scale_fill_manual(""
                    , values = cbPalette
                    , labels = c("Sentinel Absent", "Sentinel Present")
                      ) +
  theme(axis.title.x = element_blank()
        , text = element_text(size = 18)
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin())
  
SENT.BARPLOT
```

##SENT Chi-Squared

```{r SENT Chi-Squared}
SENT.CHISQ<-lapply(SENT[,c("GENERALIZED_ENVIRONMENT"
                      , "GROUP_SIZE"
                      , "DISTURBANCE_FREQUENCY"
                      )
                              ]
                         , function(x) CrossTable(x,SENT$SENTINEL_PRESENCE
                                                  , fisher = T
                                                  , chisq = T
                                                  , expected = T
                                                  , prop.c = F
                                                  , prop.t = F
                                                  , prop.chisq = F
                                                  , sresid = T
                                                  , format = 'SPSS')
                         )

SENT.CHISQ
```

##SENT Plots - Continuous Variables

```{r SENT Frequency plots}
SENT.FREQ <- SENT %>% 
  rename("Temperature (deg. C)" = "TEMPERATURE"
         , "Decimal Time" = "DECIMAL_TIME"
         , "Disturbance Frequency (per min)" = "TOTAL_FREQUENCY_OF_DISTURBANCES"
         , "Number of Crows" = "NUMBER_OF_CROWS_RECORDED") %>%
  pivot_longer(., cols = c(#"Temperature (deg. C)"
                           # , "Decimal Time"
                           "Number of Crows"
                           , "Disturbance Frequency (per min)")
               , names_to = "Var"
               , values_to = "Val") %>%
  ggplot(aes(
    x = Val
    , colour = SENTINEL_PRESENCE))+
  geom_freqpoly(linewidth = 1
                , alpha=I(.6)) +
  scale_colour_manual(values = cbPalette
                      , name = ""
                      , labels = c("Sentinel Absent", "Sentinel Present")) +
  facet_wrap(~Var, scale = "free_x") +
  theme_classic() +
  theme(legend.position = "bottom"
        , axis.title.x = element_blank()
        , text = element_text(size = 18))

SENT.FREQ
```