---
title: "PTWY"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---
```{r Libraries, include = F}
library("MANOVA.RM")
library("dplyr")
library("ggplot2")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

```{r Open DATA.SR}
PTWY.raw<-read.csv("DATA.SR.csv", stringsAsFactors = T)
str(PTWY.raw)
```

Nice, it opened as it should. Let's subset the dataset to only contain what we want.

```{r Subsetting PTWY}
PTWY<- PTWY.raw %>%
  dplyr::select(VIDEO_ID.,ID., DECIMAL_TIME, GENERALIZED_ENVIRONMENT, SENTINEL_PRESENCE, BAIT_PRESENCE, NUMBER_OF_CROWS_RECORDED, GROUP_SIZE, TOTAL_FREQUENCY_OF_DISTURBANCES, DISTURBANCE_FREQUENCY | starts_with("PTWY"))

```

Next, let's group together the transitions to vigilance, and transitions to vulnerability.
 
```{r PTWY group}
Vulnerable<-cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)
Vigilant<-cbind(PTWY$PTWY_HD.HU, PTWY$PTWY_HDP.HU)
# This does not work

```

Let's try fitting the model.

## Vulnerable

```{r PTWY MANOVA Vulnerable}
PTWY.vuln<-MANOVA.wide(cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, data = PTWY, subject = "ID.", iter=1000)
summary(PTWY.vuln)
```

Interesting... While Sentinel presence and the interaction of sentinel presence and generalized environment had no significant effects, generalized environment had an effect on the frequency of transitions to vulnerability.

**Figure out if you need to recalculate PTWY data to only include those 4 transitions. **

Let's perform a pairwise comparison.

### Post Hoc

```{r PTWY CR Vuln GENV}
simCI(PTWY.vuln, contrast = "pairwise", type = "Tukey", interaction  = F, factor = "GENERALIZED_ENVIRONMENT")

```

HMMMM... Very weird... Now its not significant... Let's compute univariate comparisons.

```{r PTWY Vulnerable univar}
PTWY.HD.HDP<-MANOVA.wide(PTWY_HD.HDP~GENERALIZED_ENVIRONMENT, data = PTWY, iter = 1000)
PTWY.HU.HD<-MANOVA.wide(PTWY_HU.HD~GENERALIZED_ENVIRONMENT, data = PTWY, iter = 1000)

p.adjust(c(PTWY.HD.HDP$resampling[,1], PTWY.HU.HD$resampling[,1]), method = "bonferroni")
#using the parametric bootstrap Wald-type statistic

```

Very cool! Turns out that only the transitions from head up to head down is significantly affected by the generalized environment. This is only if we use the Wald-type statistic and not the modified ANOVA-type statistic.

### Plots

```{r PTWY plots}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
G.ENV.PTWY<-PTWY %>%
  summarySE(measurevar = "PTWY_HU.HD"
            , groupvars = "GENERALIZED_ENVIRONMENT") %>%
    ggplot(aes(x = GENERALIZED_ENVIRONMENT
                     , y = PTWY_HU.HD
                     , color = GENERALIZED_ENVIRONMENT))+
  geom_point(position = position_dodge(width=0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(PTWY_HU.HD-se)
                    , ymax=(PTWY_HU.HD+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  theme_classic() +
  ylab("Head Up to Head Down (%)") +
  scale_colour_manual(values = cbPalette
                      , guide="none")+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin())

G.ENV.PTWY

```

## Vigilance

```{r MANOVA Vigilance}
PTWY.vigil<-MANOVA.wide(cbind(PTWY$PTWY_HD.HU, PTWY$PTWY_HDP.HU)~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, data = PTWY, subject = "ID.", iter=1000)
summary(PTWY.vigil)
```

The results of this test suggest that no transitions to vigilance are significantly affected by either the presence of a sentinel or the generalized environment.

**Figure out what to do with factors such as bait presence, group size and disturbance frequency**

## The other variables - Vulnerability

We still have bait presence, group size and disturbance frequency to look at. I want to try and make a quick function to run all the tests.

```{r PTWY Vulnerable function}
PTWY.vuln.all<-lapply(PTWY[,c("SENTINEL_PRESENCE"
                              ,"GENERALIZED_ENVIRONMENT"
                              , "GROUP_SIZE"
                              , "BAIT_PRESENCE"
                              , "DISTURBANCE_FREQUENCY")
                              ]
                         , function(x) MANOVA.wide(cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)~x, data = PTWY, iter = 1000)
                         )
```
```{r PTWY Vulnerable summaries, echo = F}
PTWY.vuln.SENT<-summary(PTWY.vuln.all$SENTINEL_PRESENCE)
PTWY.vuln.GENV<-summary(PTWY.vuln.all$GENERALIZED_ENVIRONMENT)
PTWY.vuln.GS<-summary(PTWY.vuln.all$GROUP_SIZE)
PTWY.vuln.BP<-summary(PTWY.vuln.all$BAIT_PRESENCE)
PTWY.vuln.DF<-summary(PTWY.vuln.all$DISTURBANCE_FREQUENCY)
```
```{r PTWY Vulnerable Results}
Factor<-c("Sentinel Presence", "Generalized Environment", "Group Size", "Bait Presence", "Disturbance Frequency")
Results.vuln<-cbind(Factor, rbind(PTWY.vuln.SENT, PTWY.vuln.GENV, PTWY.vuln.GS, PTWY.vuln.BP, PTWY.vuln.DF))
Results.vuln
```

As seen above, Gen. env has a  significant effect (p-value = 0.007). Interestingly,  group size (p-value ~ 0.072), bait presence (p-value ~ 0.06), and disturbance frequency (p-value ~ 0.01) are have either significant or marginally significant effects on the frequency of pathways to vulnerability.

### Post Hoc - Disturbance Frequency

```{r PTWY CR Vuln DF}
PTWY.DF.Vuln<-MANOVA.wide(cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)
simCI(PTWY.DF.Vuln, contrast = "pairwise", type = "Tukey", interaction  = F, factor = "DISTURBANCE_FREQUENCY")

```

HMMMMM... Not significant... Univariate analysis time.

```{r PTWY.DF Vulnerable univar}
PTWY.DF.HD.HDP<-MANOVA.wide(PTWY_HD.HDP~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)
PTWY.DF.HU.HD<-MANOVA.wide(PTWY_HU.HD~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)

p.adjust(c(PTWY.DF.HD.HDP$resampling[,1], PTWY.DF.HU.HD$resampling[,1]), method = "bonferroni")
#using the parametric bootstrap Wald-type statistic

```

Curious... the frequency of the pathway HD>HDPeck is marginally significant. Let's quickly plot it.

### Plots

```{r PTWY.DF.vuln plots}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
DF.PTWY.vuln<-PTWY %>%
  summarySE(measurevar = "PTWY_HD.HDP"
            , groupvars = "DISTURBANCE_FREQUENCY") %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
                     , y = PTWY_HD.HDP
                     , color = DISTURBANCE_FREQUENCY))+
  geom_point(position = position_dodge(width=0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(PTWY_HD.HDP-se)
                    , ymax=(PTWY_HD.HDP+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  theme_classic() +
  ylab("Head Down to Head Down + Peck (%)") +
  scale_colour_manual(values = cbPalette
                      , guide="none")+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()) +
  scale_x_discrete(limits = c("LOW", "MEDIUM", "HIGH"))

DF.PTWY.vuln

```

THATS UNEXPECTED... well.. not that unexpected. It mirrors the results of the peck data! As disturbance frequency increases, so does the peck rate, and therefore the frequency of head down to head down + peck. It could be a sense of urgency.

## The other variables - Vulnerability

We still have bait presence, group size and disturbance frequency to look at. I want to try and make a quick function to run all the tests.

```{r PTWY Vigilant function}
PTWY.vigil.all<-lapply(PTWY[,c("SENTINEL_PRESENCE"
                              ,"GENERALIZED_ENVIRONMENT"
                              , "GROUP_SIZE"
                              , "BAIT_PRESENCE"
                              , "DISTURBANCE_FREQUENCY")
                              ]
                         , function(x) MANOVA.wide(cbind(PTWY$PTWY_HD.HU, PTWY$PTWY_HDP.HU)~x, data = PTWY, iter = 1000)
                         )
```
```{r PTWY Vigilant summaries, echo = F}
PTWY.vigil.SENT<-summary(PTWY.vigil.all$SENTINEL_PRESENCE)
PTWY.vigil.GENV<-summary(PTWY.vigil.all$GENERALIZED_ENVIRONMENT)
PTWY.vigil.GS<-summary(PTWY.vigil.all$GROUP_SIZE)
PTWY.vigil.BP<-summary(PTWY.vigil.all$BAIT_PRESENCE)
PTWY.vigil.DF<-summary(PTWY.vigil.all$DISTURBANCE_FREQUENCY)
```
```{r PTWY Vigilant Results}
Results.vigil<-cbind(Factor, rbind(PTWY.vigil.SENT, PTWY.vigil.GENV, PTWY.vigil.GS, PTWY.vigil.BP, PTWY.vigil.DF))
Results.vigil
```

Another interesting result! Disturbance frequency is marginally significant for transitions to vigilance. Same workflow as above.

```{r PTWY DF CR Vigil}
PTWY.DF.Vigil<-MANOVA.wide(cbind(PTWY$PTWY_HD.HU, PTWY$PTWY_HDP.HU)~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)
simCI(PTWY.DF.Vigil, contrast = "pairwise", type = "Tukey", interaction  = F, factor = "DISTURBANCE_FREQUENCY")

```

Pairwise comparisons for medium - high is marginally significant.

```{r PTWY.DF Vigil univar}
PTWY.DF.HD.HU<-MANOVA.wide(PTWY_HD.HU~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)
PTWY.DF.HDP.HU<-MANOVA.wide(PTWY_HDP.HU~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)

p.adjust(c(PTWY.DF.HD.HU$resampling[,1], PTWY.DF.HDP.HU$resampling[,1]), method = "bonferroni")
#using the parametric bootstrap Wald-type statistic

```

Disturbance frequency has a marginally significant effect on transitions from Head Down + Peck to Head Up. Plot time!

```{r PTWY.DF.vigil plots}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
DF.PTWY.vigil<-PTWY %>%
  summarySE(measurevar = "PTWY_HDP.HU"
            , groupvars = "DISTURBANCE_FREQUENCY") %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
                     , y = PTWY_HDP.HU
                     , color = DISTURBANCE_FREQUENCY))+
  geom_point(position = position_dodge(width=0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(PTWY_HDP.HU-se)
                    , ymax=(PTWY_HDP.HU+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  theme_classic() +
  ylab("Head Down + Peck to Head Up  (%)") +
  scale_colour_manual(values = cbPalette
                      , guide="none")+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()) +
  scale_x_discrete(limits = c("LOW", "MEDIUM", "HIGH"))

DF.PTWY.vigil

```

# Frequency

We'll be fitting another manova on the frequency data.

```{r Subsetting FREQ}
FREQ<- PTWY.raw %>%
  dplyr::select(VIDEO_ID.,ID., DECIMAL_TIME, GENERALIZED_ENVIRONMENT, SENTINEL_PRESENCE, BAIT_PRESENCE, NUMBER_OF_CROWS_RECORDED, GROUP_SIZE, TOTAL_FREQUENCY_OF_DISTURBANCES, DISTURBANCE_FREQUENCY | starts_with("FREQ"))

```

Let's try fitting the model.

## Vulnerable

```{r FREQ MANOVA Vulnerable}
FREQ.vuln<-MANOVA.wide(cbind(FREQ$FREQ_HD.HDP, FREQ$FREQ_HU.HD)~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, data = FREQ, subject = "ID.", iter=1000)
summary(FREQ.vuln)
```

Same as with the PTWY data, only the generalized environment had a significant effect, and that only with the Wald-type statistic. Weird... 

### Post Hoc

```{r FREQ CR Vuln}
simCI(FREQ.vuln, contrast = "pairwise", type = "Tukey", interaction  = F, factor = "GENERALIZED_ENVIRONMENT")

```

Womp womp. Not significant. :/
However, this still follows the same pattern as the PTWY analysis. Let's perform a univariate analysis to confirm.

```{r FREQ Vulnerable univar}
FREQ.HD.HDP<-MANOVA.wide(FREQ_HD.HDP~GENERALIZED_ENVIRONMENT, data = FREQ, iter = 1000)
FREQ.HU.HD<-MANOVA.wide(FREQ_HU.HD~GENERALIZED_ENVIRONMENT, data = FREQ, iter = 1000)

p.adjust(c(FREQ.HD.HDP$resampling[,1], FREQ.HU.HD$resampling[,1]), method = "bonferroni")
#using the parametric bootstrap Wald-type statistic

```

Cool! HU>HD is marginally insignificant with a p-value of ~0.08.
Onto vigilance!



<!-- # Old code below -->
<!-- # ```{r PTWY Analysis} -->
<!-- # PTWY.ENV <- '  -->
<!-- #  -->
<!-- # # measurement model  -->
<!-- #  -->
<!-- # ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TOTAL_FREQUENCY_OF_DISTURBANCES  -->
<!-- #  -->
<!-- # # intercepts  -->
<!-- #  -->
<!-- # CODED_ENV ~ 1 -->
<!-- # CODED_BAIT ~ 1 -->
<!-- # DECIMAL_TIME ~ 1 -->
<!-- # TOTAL_FREQUENCY_OF_DISTURBANCES ~ 1 -->
<!-- # '  -->
<!-- #  -->
<!-- # PTWY.Fit.ENV <- sem(PTWY.ENV, data = PTWY)  -->
<!-- #  -->
<!-- # summary(PTWY.Fit.ENV, fit.measures = TRUE, standardized = TRUE) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r PTWY.SOC Analysis} -->
<!-- # PTWY.Mod.SOC <- '  -->
<!-- #  -->
<!-- # # measurement model  -->
<!-- # SOCIAL =~ CODED_GROUP_SIZE  -->
<!-- #  -->
<!-- # # intercepts -->
<!-- # CODED_GROUP_SIZE ~ 1 -->
<!-- # '  -->
<!-- #  -->
<!-- # PTWY.Fit.SOC <- sem(PTWY.Mod.SOC, data = PTWY)  -->
<!-- #  -->
<!-- # summary(PTWY.Fit.SOC, fit.measures = TRUE, standardized = TRUE) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r PTWY Analysis} -->
<!-- # PTWY.Mod <- '  -->
<!-- #  -->
<!-- # # measurement model  -->
<!-- #  -->
<!-- # ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TEMPERATURE + TOTAL_FREQUENCY_OF_DISTURBANCES  -->
<!-- #  -->
<!-- # VIGILANCE =~ PTWY_HDP.HU + PTWY_HD.HU + PTWY_M.HU  -->
<!-- #  -->
<!-- # FORAGING =~ PTWY_HD.HDP + PTWY_HU.HD + PTWY_M.HD  -->
<!-- #  -->
<!-- #   -->
<!-- #  -->
<!-- # # regressions  -->
<!-- #  -->
<!-- # VIGILANCE ~ ENVIRONMENT + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE -->
<!-- #  -->
<!-- # FORAGING ~ ENVIRONMENT + VIGILANCE + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE -->
<!-- #  -->
<!-- # '  -->
<!-- #  -->
<!-- # PTWY.Fit <- sem(PTWY.Mod, data = PTWY)  -->
<!-- #  -->
<!-- # summary(PTWY.Fit, fit.measures = TRUE, standardized = TRUE) -->
<!-- # ``` -->